### 2021-03-19 ch/4-structure.tex
Nearly there

\chapter{Summary of contributions}\label{ch:content}
The following is a summary of the main contributions in this thesis.
All presented approaches deal with dataset bias that is closely linked to a special attribute $s$.
Furthermore, the approaches all rely on some form of side information with allows us to overcome dataset bias.
This side information is always significantly easier to obtain than unbiased data.

@@

In order to construct the target labels,
we use side information about summary statistics for an balanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\).
These target rates are the only parameters of the model.
The target labels represent an uncertain estimate of labels corresponding to a balanced dataset.
Via the sum rule of probabilities, it is possible to express the model likelihood in terms of these target labels,
such that maximizing the likelihood corresponds to improving the prediction of the target labels
instead of the given (biased) training labels.

@@

As an example, consider the scenario where someone wants to create a classifier
to distuinguish between sheep and cows, that is supposed to work anywhere on earth.
However, they take a shortcut while creating the dataset
and take all their sheep images from hot and dry countries
and all their cow images from mild and rainy countries.
In this case, the dataset is lacking cow images in dry landscapes,
and is lacking sheep from green landscapes;
the dataset exhibits a strong sampling bias.
The result is that even though the labels correctly correspond to cows and sheep,
they do not point reliably to the right target anymore.
As background colour is easier to recognise with a \ac{CNN} than animal species,
the labels have effectively been turned into landscape labels.
In other words, landscape has become a \emph{spurious attribute}.
%
% As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
% Say, we initially have a very diverse and balanced training set,
% but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
% and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
% The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
% As hair colour is easier to recognise with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
% In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).
However, there is a difference in emphasis between a \emph{sensitive} and a \emph{spurious} attribute:
the former indicates that the attribute should not be used for legal or ethical reasons,
whereas the latter can be any attribute that is associated with the class label \(y\) in an undesired way
that leads to lower quality predictions.

The method, proposed in the previous paper (\chapref{ch:paper1}),
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colours to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair colour groups given by \(s\).
This is the problem tackled in the previous paper,
and yet, if we were to equalize the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

@@

\section{Overcoming sampling bias with unlabelled deployment set}\label{sec:zsf}
A shortcoming of the approach from the previous publication \citep{kehrenberg2020nullsampling}
is its reliance on a representative set which has labels for the spurious attbiute \(s\).
As discussed, it is necessary to make use of \emph{some} kind of side information,
but perhaps we can relax some requirements.
In particular, while it is already easier to collect data without \(y\) labels (but with \(s\) labels),
it is even easier to collect data without \emph{any} label.
Thus, requiring only an unlabelled context set would improve the applicability of the method.
\citet{kehrenberg2020zeroshot} (\chapref{ch:paper3}) presents an approach based on that idea.

@@

The distribution matching is realized with adversarial networks which compare batches of samples,
in order to try to distinguish data drawn from the training set and the deployment set.
This process requires balanced batches as an inductive bias,
because the network will only learn the intended difference between training and deployment set,
if the drawn batches exhibit this difference.
For example, if batches drawn from Coloured MNIST differ not in colour but in digit class,
then distribution matching will learn to change digit shapes.
Balancing batches from the training set is easily possible with the available labels,
but those are not available for the deployment set,
so we use clustering techniques to identify the different groups in the deployment set,
and then draw samples for the batches at equal rate from all clusters.
It is important to note here that imperfections in the clustering are not a problem
as long as the show on average the intended difference between training and deployment set.

>>>>

\chapter{Summary of contributions}\label{ch:content}
The following is a summary of the main contributions in this thesis.
All presented approaches deal with dataset bias that is closely linked to a special attribute $s$.
In terms of \citet{kamishima2012fairness}'s taxonomy (see section \secref{fair-classifiers}),
this can be described as tackling \emph{negative legacy} that is mediated by \emph{prejudice}.
Furthermore, the approaches all rely on some form of side information with allows us to overcome dataset bias.
This side information is always significantly easier to obtain than unbiased data.

@@

In order to construct the target labels,
we use side information about summary statistics for an balanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\).
These target rates are the only parameters of the model.
The target labels represent an uncertain estimate of labels corresponding to a balanced dataset.
Via the sum rule of probabilities, it is possible to express the model likelihood in terms of these target labels,
such that maximising the likelihood corresponds to improving the prediction of the target labels
instead of the given (biased) training labels.

@@

As an example, consider the scenario where someone wants to create a classifier
to distinguish between sheep and cows, that is supposed to work anywhere on earth.
However, they take a shortcut while creating the dataset
and take all their sheep images from hot and dry countries
and all their cow images from mild and rainy countries.
In this case, the dataset is lacking cow images in dry landscapes,
and is lacking sheep from green landscapes;
the dataset exhibits a strong sampling bias.
The result is that even though the labels correctly correspond to cows and sheep,
they do not point reliably to the right target anymore.
As background colour is easier to recognise with a \ac{CNN} than animal species,
the labels have effectively been turned into landscape labels.
In other words, landscape has become a \emph{spurious attribute}.
%
% As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
% Say, we initially have a very diverse and balanced training set,
% but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
% and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
% The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
% As hair colour is easier to recognise with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
% In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).
However, there is a difference in emphasis between a \emph{sensitive} and a \emph{spurious} attribute:
the former indicates that the attribute should not be used for legal or ethical reasons,
whereas the latter can be any attribute that is associated with the class label \(y\) in an undesired way
that leads to lower quality predictions.

The method, proposed in the previous paper (\chapref{ch:paper1}),
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colours to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair colour groups given by \(s\).
This is the problem tackled in the previous paper,
and yet, if we were to equalise the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

@@

\section{Overcoming sampling bias with unlabelled deployment set}\label{sec:zsf}
A shortcoming of the approach from the previous publication \citep{kehrenberg2020nullsampling}
is its reliance on a representative set which has labels for the spurious attribute \(s\).
As discussed, it is necessary to make use of \emph{some} kind of side information,
but perhaps we can relax some requirements.
In particular, while it is already easier to collect data without \(y\) labels (but with \(s\) labels),
it is even easier to collect data without \emph{any} label.
Thus, requiring only an unlabelled context set would improve the applicability of the method.
\citet{kehrenberg2020zeroshot} (\chapref{ch:paper3}) presents an approach based on that idea.

@@

The distribution matching is realised with adversarial networks which compare batches of samples,
in order to try to distinguish data drawn from the training set and the deployment set.
This process requires balanced batches as an inductive bias,
because the network will only learn the intended difference between training and deployment set,
if the drawn batches exhibit this difference.
For example, if batches drawn from Coloured MNIST differ not in colour but in digit class,
then distribution matching will learn to change digit shapes.
Balancing batches from the training set is easily possible with the available labels,
but those are not available for the deployment set,
so we use clustering techniques to identify the different groups in the deployment set,
and then draw samples for the batches at equal rate from all clusters.
It is important to note here that imperfections in the clustering are not a problem
as long as the show on average the intended difference between training and deployment set.

### 2021-03-19 ch/6-discussion.tex
Nearly there

While this by no means covers \emph{all} possible biases,
it contributes to a growing literature that tries to tackle this problem.
One could ask if there is one method that is able to cover all possible dataset biases,
but I think there is a strong argument to be made that no general method can exist,
because it is, \eg, not possible to describe in general what is spurious information and what is relevant information.
Nevertheless, finding methods that are more generally applicable is a worthy goal.

Another potential extension of this work is to work with other modalities.
The experiments were all performed on either tabular or image data
--- the reason predominatly being ease of visualization ---
so working with audio (especially speech) or text could present new interesting challenges.

Furthermore, as it currently stands,
the \ac{ML} practitioner has to know the bias in the data very well in order to choose a method to correct it.
It would be desirable to have a simple algorithm for deciding which method to use.

Thus, correcting dataset bias remains a challenging topic
and one that is relevant to today's machine learning applications.
Any cutting-edge \ac{ML} system will have to deal with imperfect data,
especially if the collected data is human-related.
The possible effects of these data imperfections are certainly highly undesirable:
a photo tagging service might only work for a certain kind of person;
a speech recognition system might only work for a certain kind of dialect.
If, in these situations, enough representative (but unlabelled) data is available,
then the methods presented here can be used to try and correct the problem.

One area of machine learning that has recently seen increased interest is unsupervised learning.
The latter two chapters also make use of it to some degree.
The exciting promise of unsupervised learning in general is
that labour-intensive labelling is not needed and so vast amounts of existing, unlabelled data can be put to good use.
One could ask the question whether bias-correcting methods are still needed, with access to so much data.
It could be that, while the data is certainly not perfect,
there is so much of it that the biased parts ``cancel each other out''.
However, recent investigations of the GPT models \citep{radford2018improving,radford2019language,brown2020language} do not seem to bear this out \citep{khalifa2021distributional}.
One reason for this might be the way these models are trained at the moment:
they maximise the probability assigned to the next token.
Thus, such a model has to account for the wide array of human opinions and assign a non-zero probability to all of them.
So, when asked to summarise a text \citep{stiennon2020learning}, GPT does not give the \emph{best} summary;
it gives a summary that an average person might have written.
However, with the help of a very high quality labelled dataset (that was expensive to create),
GPT could be finetuned to actually produce very good summaries.
I suspect this pattern of learning the basics in an unsupervised fashion
and then finetuning with high-quality labels will continue in the future.

>>>>

While this by no means covers \emph{all} possible biases,
it contributes to a growing literature that tries to tackle this problem.
One could ask if there is one method that is able to cover all possible dataset biases,
but I think there is a strong argument to be made that no general method can exist,
because it is, \eg, not possible to describe in general what is spurious information and what is relevant information.
Nevertheless, finding methods that are more generally applicable is a worthy goal.
One immediate avenue for future work is
the combination of label bias correction and sampling bias correction.
For example, it might be possible to leverage an unlabelled context set for correcting label bias as well.

There are also certain technical improvements that could be made.
In general, it would be desirable to have more theoretical bounds on performance.
This likely involves specifying the requirements for the algorithms more precisely.
% which is also useful information for potential users.
The method based on target labels would furthermore benefit from better-calibrated probability outputs.
\acf{GP} classifiers show generally good calibration,
but cannot be easily applied in domains where deep neural networks are used.
On the other hand, neural networks are known to be overconfident
(especially when using ReLU activations \citep{hein2019relu}),
so to apply the proposed method to images directly is not straightforward.
The other two proposed methods would benefit from improved adversarial training procedures.
Training these models is not straightforward as the losses have to be balanced
and in some cases, the update schedule needs to be changed.
Both the calibration of neural networks and the stability of adversarial training
are issues that are widely recognised in the \ac{ML} community,
so there is hope that progress will be made on these.

Another potential extension of this work is to work with other modalities.
The experiments were all performed on either tabular or image data
--- the reason predominantly being ease of visualisation ---
so working with audio (especially speech) or text could present new interesting challenges.

Furthermore, as it currently stands,
the \ac{ML} practitioner has to know the bias in the data very well in order to choose a method to correct it.
It would be desirable to have a simple algorithm for deciding which method to use.
A related limitation is that the dataset bias considered in this thesis
is assumed to be linked to a special attribute \(s\).
While \(s\) can be very high-dimensional and is not limited to, say, binary attributes,
this nevertheless represents a restriction
that excludes large areas of dataset biases.

In any case, correcting dataset bias remains a challenging topic
and one that is relevant to today's machine learning applications.
Any cutting-edge \ac{ML} system will have to deal with imperfect data,
especially if the collected data is human-related.
The possible effects of these data imperfections are certainly highly undesirable:
a photo tagging service might only work for a certain kind of person;
a speech recognition system might only work for a certain kind of dialect.
If, in these situations, sufficient representative (but unlabelled) data is available,
then the methods presented here can be used to try and correct the problem.

One area of machine learning that has recently seen increased interest is unsupervised learning.
The latter two chapters also make use of it to some degree.
The exciting promise of unsupervised learning in general is
that labour-intensive labelling is not needed and so vast amounts of existing, unlabelled data can be put to good use.
One could ask the question whether bias-correcting methods are still needed, with access to so much data.
It could be that, while the data is certainly not perfect,
there is so much of it that the biased parts ``cancel each other out''.
However, recent investigations of the GPT models \citep{radford2018improving,radford2019language,brown2020language} do not seem to bear this out \citep{khalifa2021distributional}.
One reason for this might be the way these models are trained at the moment:
they maximise the probability assigned to the next token.
Thus, such a model has to account for the wide array of human opinions and assign a non-zero probability to all of them.
So, when asked to summarise a text \citep{stiennon2020learning}, GPT does not give the \emph{best} summary;
it gives a summary that an average person might have written.
However, with the help of a very high-quality labelled dataset (that was expensive to create),
GPT could be finetuned to actually produce very good summaries.
I suspect this pattern of learning the basics in an unsupervised fashion
and then finetuning with high-quality labels will continue in the future.
With these massive models,
it is, more than ever, crucial to build tools to make the biases within the models transparent.
Given the black-box nature of neural networks, this represents a major challenge.
% When examining these models trained on massive datasets through the lens of the dataset biases discussed in \chapref{ch:related-work},
% we encounter a novel problem:
% somewhere in the model,
% (an approximation of) the fundamental structures of reality are likely represented,
% but we cannot easily access this.

Finally, an important issue is the communication between human and machine.
The methods presented in this thesis have all strived to make it easy for human operators to understand what is going on:
the method in \chapref{ch:paper1} is configured with simple statistics;
the method in \chapref{ch:paper2} produces invariant images that can be inspected;
and the method in \chapref{ch:paper3} has the same capability
(though it is not part of the core functionality).
However, this is only a beginning.
It is still not easy to notice that a given dataset has problems,
and machines on their own cannot notice the problem,
because they do not (yet) have a complete understanding about what it is that humans care about.
Human preferences are incredibly complex and hard to predict for today's algorithms.
Thus, it is important that machines get feedback early and often,
to keep them aligned with human goals.
(The aforementioned \citet{stiennon2020learning} is arguably an example of this.)
Applied to the problem of dataset bias,
this could mean visualising correlations in the dataset,
or routinely producing invariant representations to show what the network thinks it is meant to learn.

### 2021-03-19 paper1/paper1.tex
Nearly there

\section{Discussion and conclusion}
Fairness is fundamentally not a challenge of algorithms alone, but very much a sociological challenge.
%
A lot of proposals have emerged recently for defining and obtaining fairness in machine learning-based decision making systems. 
%
% A vast amount
The vast majority of academic work has focused on two categories of definitions: statistical (group) notions of fairness and individual notions of fairness (see~\citet{verma2018fairness} for at least twenty different notions of fairness).
%
Statistical notions are easy to verify but do not provide protections to individuals. 
%
Individual notions do give individual protections but need strong assumptions, such as the availability of an agreed-upon similarity metric, which can be difficult in practice.
%
We acknowledge that a proper solution to algorithmic fairness cannot rely on statistics alone.
%
Nevertheless, these statistical fairness definitions can be helpful in understanding the problem and working towards solutions.
%
To facilitate this, at every step, the trade-offs that are present should be made very clear
and long-term effects have to be considered as well~\citep{liu2018delayed,kallus2018residual}.
% For example, it can be the case that false positives and false negatives are not equally costly,
% in which case a closer inspection of

@@

\subsection{Finding minimal $t_s$}
We express $t_s$ in terms of $\mathit{PR}_b^s$ and $\mathit{PR}_t$.
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr2}
\end{align}
Without loss of generality, we assume $\mathit{PR}_b^0<\mathit{PR}_b^1$.
As mentioned in the main text, 
$\mathit{PR}_t$ should be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$ to minimise both $t_s$.
If that is the case, then we get
\begin{align}
  t_{s=0} &= \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^0}{1 - \mathit{PR}_t}\\
  t_{s=1} &= \frac{1}{2}\frac{\mathit{PR}_b^1 - \mathit{PR}_t}{\mathit{PR}_t}~.
\end{align}
If we further assume $\mathit{PR}_b^1<\tfrac{1}{2}$,
then we also have $\mathit{PR}_t<\tfrac{1}{2}$ and thus $\mathit{PR}_t<1-\mathit{PR}_t$.
This implies that the denominator of $t_{s=1}$ is smaller and that, in turn, $t_{s=1}$ grows faster.
This faster growth means that when minimizing $t_{s=0} + t_{s=1}$, we have to concentrate on $t_{s=1}$.
The minimum is then such that $t_{s=1}$ is 0, i.e.\ $\mathit{PR}_t=\mathit{PR}_b^1$.

>>>>

\section{Discussion and conclusion}\label{sec:paper1-discussion}
Fairness is fundamentally not a challenge of algorithms alone, but very much a sociological challenge.
%
A lot of proposals have emerged recently for defining and obtaining fairness in machine learning-based decision making systems. 
%
% A vast amount
The vast majority of academic work has focused on two categories of definitions: statistical (group) notions of fairness and individual notions of fairness (see~\citet{verma2018fairness} for at least twenty different notions of fairness).
%
Statistical notions are easy to verify but do not provide protections to individuals. 
%
Individual notions do give individual protections but need strong assumptions, such as the availability of an agreed-upon similarity metric, which can be difficult in practice.
%
We acknowledge that a proper solution to algorithmic fairness cannot rely on statistics alone.
%
Nevertheless, these statistical fairness definitions can be helpful in understanding the problem and working towards solutions.
%
To facilitate this, at every step, the trade-offs that are present should be made very clear
and long-term effects have to be considered as well~\citep{liu2018delayed,kallus2018residual}.
% For example, it can be the case that false positives and false negatives are not equally costly,
% in which case a closer inspection of

@@

\subsection{Finding minimal $t_s$}
We express $t_s$ in terms of $\mathit{PR}_b^s$ and $\mathit{PR}_t$.
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr2}
\end{align}
Without loss of generality, we assume $\mathit{PR}_b^0<\mathit{PR}_b^1$.
As mentioned in the main text, 
$\mathit{PR}_t$ should be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$ to minimise both $t_s$.
If that is the case, then we get
\begin{align}
  t_{s=0} &= \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^0}{1 - \mathit{PR}_t}\\
  t_{s=1} &= \frac{1}{2}\frac{\mathit{PR}_b^1 - \mathit{PR}_t}{\mathit{PR}_t}~.
\end{align}
If we further assume $\mathit{PR}_b^1<\tfrac{1}{2}$,
then we also have $\mathit{PR}_t<\tfrac{1}{2}$ and thus $\mathit{PR}_t<1-\mathit{PR}_t$.
This implies that the denominator of $t_{s=1}$ is smaller and that, in turn, $t_{s=1}$ grows faster.
This faster growth means that when minimising $t_{s=0} + t_{s=1}$, we have to concentrate on $t_{s=1}$.
The minimum is then such that $t_{s=1}$ is 0, i.e.\ $\mathit{PR}_t=\mathit{PR}_b^1$.

### 2021-03-19 paper3/paper3.tex
Nearly there

%% Recommended, but optional, packages for figures and better typesetting:
%\usepackage{microtype}
%\usepackage{graphicx}
%\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{booktabs} % for professional tables
%\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsthm}
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{xcolor}         %for coloring the text on colorMNIST
%\usepackage{comment}
%\usepackage{enumitem}
%\newtheorem{prop}{Proposition}
%\newcommand{\Xcal}{\mathcal{X}}
%\newcommand{\Bcal}{\mathcal{B}}
%\newcommand{\Ycal}{\mathcal{Y}}
%\newcommand{\Scal}{\mathcal{S}}
%\newcommand{\Lcal}{\mathcal{L}}
%\newcommand{\Dcal}{\mathcal{D}}
%\newcommand{\ie}{i.\,e.}
%\newcommand{\Ie}{I.\,e.}
%\newcommand{\eg}{e.\,g.}
%\newcommand{\Eg}{E.\,g.}

@@

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine (e.g. \citep{DunYiLanReetal19}), and socially important contexts such as the allocation of healthcare, education, and credit (\eg\ \citep{HurAde17,RagBarKleLev20}). 
%
Efficiency can be improved, costs can be reduced, and personalization of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/celeba_dep_recon.png}
    \caption{%
    Visualization of our method's solutions for the CelebA dataset, with``smiling females'' as the missing subgroup.
    Column 1 shows the original images from $x$ from the deployment set of CelebA.
    Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
    Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
    Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
    }%
    \label{fig:celeba-recons}
\end{figure}

@@

A limitation of the presented approach is that, for constructing the perfect bags used to train the disentangling algorithm, we have relied on knowing the number of clusters \emph{a priori}, something that, in practice, is perhaps not the case. 
%
Removing this dependency through automatic determination of the number of clusters would generalise our method further but this line of research is challenging and extends beyond the scope of the current paper. 
%
One difficulty is that we need to ensure that the small but salient clusters are correctly identified. 
%
% In the case of 2-digit Colored MNIST, for example, the deployment set may contain only a small portion of {\color{purple}purple} fours relative to the other subgroups, meaning that the cluster they form can be easily overlooked by a clustering algorithm in favor of larger but less salient clusters (which may be sub-clusters of other digit/color combinations). 
The cluster formed by an underrepresented subgroup can be easily overlooked by a clustering algorithm in favor of larger but less salient clusters
(which may be sub-clusters of other, larger, subgroups).

@@

% In this paper, we aim to highlight the problem of zero-label stratification that can result from systematic bias, and to make a case for it as something to be seriously considered (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
%Clearly, the use of algorithmic approaches still requires human review in a manner that is similar to error auditing, but is less dependent on the specific human auditor to initially identify the stratifica- tion. While algorithmic approaches to measurement can reduce burden on human analysts and take advantage of learned encodings to identify subclasses, their efficacy is limited by the difficulty of separating important subclasses in the feature space analysed.

@@

To validate the disentangling step, we compare with three other baselines.
In all cases, the training set is balanced in the same fashion required for our method.
This is similar in effect to the sampling method proposed by \citet{kamiran2012data} but with those samples with the same class label as the missing sources upsampled, based on knowledge of $\text{dim}(\mathcal{S})$.
We denote a classifier trained with cross-entropy loss on this data as \texttt{ERM}.
The second of the aforementioned baselines is \texttt{DRO} \citep{HasSriNamLia18}, which functions without subgroup labels by minimizing the worst-case training loss over all possible groups that are above a certain minimum size;
along with a variant of it, \texttt{gDRO} \citep{sagawa2019distributionally}, that exploits subgroup label information but is then only applicable when $\text{dim}(\mathcal{S}_{tr}) > 1$.
Third, we have the \texttt{LfF} model proposed by \citet{nam2020learning} that works by reweighting the cross-entropy loss of a classifier using the predictions of a purposely biased sister network.

%
%%%%%OR
%During training, the {\color{green}green} color group is observed in both digits (sources $\Omega_{y=\text{'two' }s=\color{green}green}$ and $\Omega_{y=\text{'four' }s=\color{green}green}$), but for the {\color{purple}purple} color group, we only observe the one-sided outcome ($\Omega_{y=\text{'two'}s=\color{purple}purple}$), so the source $\Omega_{y=\text{'four'}s=\color{green}green}$ defines invisible demographic.  
%
\subsection{Coloured MNIST}
The MNIST dataset \citep{lecun1998gradient} consists of 70,000 (60,000 designated for training, 10,000 for testing) images of grey-scale hand-written digits. We colour the digits following a similar procedure to that outlined by \citep{kim2019learning}, randomly assigning each sample one of ten distinct RGB colours. Each source is then a combination of digit-class (class label) and colour (subgroup label). We use no data-augmentation aside from symmetrically zero-padding the images to be of size 32x32.
We create imbalance in both $D_{dep}$ and $D_{tr}$ by sub-sampling the remaining sources;
we do this to render a more realistic setting in which the deployment set is not innately balanced and demands preliminary clustering to construct approximately perfect bags. The sub-sampling proportions used for each set of experiments can be found in the appendix.

@@

Fig.~\ref{fig:cmnist-2v4-miss-s} shows that the problem of \emph{missing subgroups} is harder to solve.
Again we see that poor clustering (in the case of \texttt{K-means}) is detrimental to the disentangling procedure, though for all balancing strategies the IQR is significantly higher than observed in the MS setting, along with there being a number of extreme outliers. The median, however, remains high, which is reflective of the ``hit-or-miss'' performance of the method under these conditions, but where the number of hits far outweighs the number of misses. Visualization of the reconstructions with $z_y$ zeroed-out suggest that misses often mostly occurred due to the semantic information being concentrated in $z_s$ while $z_y$ is left to contain only residual information, even when $z_s$ was set to be one-dimensional and binarised. We leave it to future work to explore how to better offset such degeneracies.

>>>>

%% Recommended, but optional, packages for figures and better typesetting:
%\usepackage{microtype}
%\usepackage{graphicx}
%\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{booktabs} % for professional tables
%\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsthm}
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{xcolor}         %for colouring the text on colourMNIST
%\usepackage{comment}
%\usepackage{enumitem}
%\newtheorem{prop}{Proposition}
%\newcommand{\Xcal}{\mathcal{X}}
%\newcommand{\Bcal}{\mathcal{B}}
%\newcommand{\Ycal}{\mathcal{Y}}
%\newcommand{\Scal}{\mathcal{S}}
%\newcommand{\Lcal}{\mathcal{L}}
%\newcommand{\Dcal}{\mathcal{D}}
%\newcommand{\ie}{i.\,e.}
%\newcommand{\Ie}{I.\,e.}
%\newcommand{\eg}{e.\,g.}
%\newcommand{\Eg}{E.\,g.}

@@

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine (e.g. \citep{DunYiLanReetal19}), and socially important contexts such as the allocation of healthcare, education, and credit (\eg\ \citep{HurAde17,RagBarKleLev20}). 
%
Efficiency can be improved, costs can be reduced, and personalisation of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/celeba_dep_recon.png}
    \caption{%
    Visualisation of our method's solutions for the CelebA dataset, with``smiling females'' as the missing subgroup.
    Column 1 shows the original images from $x$ from the deployment set of CelebA.
    Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
    Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
    Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
    }%
    \label{fig:celeba-recons}
\end{figure}

@@

A limitation of the presented approach is that, for constructing the perfect bags used to train the disentangling algorithm, we have relied on knowing the number of clusters \emph{a priori}, something that, in practice, is perhaps not the case. 
%
Removing this dependency through automatic determination of the number of clusters would generalise our method further but this line of research is challenging and extends beyond the scope of the current paper. 
%
One difficulty is that we need to ensure that the small but salient clusters are correctly identified. 
%
% In the case of 2-digit Coloured MNIST, for example, the deployment set may contain only a small portion of {\color{purple}purple} fours relative to the other subgroups, meaning that the cluster they form can be easily overlooked by a clustering algorithm in favour of larger but less salient clusters (which may be sub-clusters of other digit/colour combinations). 
The cluster formed by an underrepresented subgroup can be easily overlooked by a clustering algorithm in favour of larger but less salient clusters
(which may be sub-clusters of other, larger, subgroups).

@@

% In this paper, we aim to highlight the problem of zero-label stratification that can result from systematic bias, and to make a case for it as something to be seriously considered (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
%Clearly, the use of algorithmic approaches still requires human review in a manner that is similar to error auditing, but is less dependent on the specific human auditor to initially identify the stratification. While algorithmic approaches to measurement can reduce burden on human analysts and take advantage of learned encodings to identify subclasses, their efficacy is limited by the difficulty of separating important subclasses in the feature space analysed.

@@

To validate the disentangling step, we compare with three other baselines.
In all cases, the training set is balanced in the same fashion required for our method.
This is similar in effect to the sampling method proposed by \citet{kamiran2012data} but with those samples with the same class label as the missing sources upsampled, based on knowledge of $\text{dim}(\mathcal{S})$.
We denote a classifier trained with cross-entropy loss on this data as \texttt{ERM}.
The second of the aforementioned baselines is \texttt{DRO} \citep{HasSriNamLia18}, which functions without subgroup labels by minimising the worst-case training loss over all possible groups that are above a certain minimum size;
along with a variant of it, \texttt{gDRO} \citep{sagawa2019distributionally}, that exploits subgroup label information but is then only applicable when $\text{dim}(\mathcal{S}_{tr}) > 1$.
Third, we have the \texttt{LfF} model proposed by \citet{nam2020learning} that works by reweighting the cross-entropy loss of a classifier using the predictions of a purposely biased sister network.

%
%%%%%OR
%During training, the {\color{green}green} colour group is observed in both digits (sources $\Omega_{y=\text{'two' }s=\color{green}green}$ and $\Omega_{y=\text{'four' }s=\color{green}green}$), but for the {\color{purple}purple} colour group, we only observe the one-sided outcome ($\Omega_{y=\text{'two'}s=\color{purple}purple}$), so the source $\Omega_{y=\text{'four'}s=\color{green}green}$ defines invisible demographic.  
%
\subsection{Coloured MNIST}
The MNIST dataset \citep{lecun1998gradient} consists of 70,000 (60,000 designated for training, 10,000 for testing) images of grey-scale hand-written digits. We colour the digits following a similar procedure to that outlined by \citep{kim2019learning}, randomly assigning each sample one of ten distinct RGB colours. Each source is then a combination of digit-class (class label) and colour (subgroup label). We use no data-augmentation aside from symmetrically zero-padding the images to be of size 32x32.
We create imbalance in both $D_{dep}$ and $D_{tr}$ by sub-sampling the remaining sources;
we do this to render a more realistic setting in which the deployment set is not innately balanced and demands preliminary clustering to construct approximately perfect bags. The sub-sampling proportions used for each set of experiments can be found in the appendix.

@@

Fig.~\ref{fig:cmnist-2v4-miss-s} shows that the problem of \emph{missing subgroups} is harder to solve.
Again we see that poor clustering (in the case of \texttt{K-means}) is detrimental to the disentangling procedure, though for all balancing strategies the IQR is significantly higher than observed in the MS setting, along with there being a number of extreme outliers. The median, however, remains high, which is reflective of the ``hit-or-miss'' performance of the method under these conditions, but where the number of hits far outweighs the number of misses. Visualisation of the reconstructions with $z_y$ zeroed-out suggest that misses often mostly occurred due to the semantic information being concentrated in $z_s$ while $z_y$ is left to contain only residual information, even when $z_s$ was set to be one-dimensional and binarised. We leave it to future work to explore how to better offset such degeneracies.
